---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: yes
    
---


## Loading and preprocessing the data

```{r}
##Please note-- I am re-taking this course so some of the commits are old
unzip("activity.zip")
act <- read.csv("activity.csv", header = TRUE)
```

## What is mean total number of steps taken per day?

```{r}
totalsteps <- tapply(act$steps, act$date, sum, na.rm = TRUE)
```  
The histogram below shows the total steps per day  
```{r}
hist(totalsteps, breaks = 50, xlab = "total steps")
```

```{r}
meansteps <- mean(totalsteps)
mediansteps <-median(totalsteps)
```
The mean total number of steps per day is `r meansteps`.  
The median total number of steps per day is `r mediansteps`.  

## What is the average daily activity pattern?

```{r}
intervalsteps <- tapply(act$steps, act$interval, mean, na.rm=TRUE)
intervals <- as.data.frame(intervalsteps)
intervals$int <- rownames(intervals)
colnames(intervals) <- c("mean", "int")
```

Here is a plot of average steps taken during each 5-minute interval, across all days
```{r}
plot(intervals$int, intervals$mean, type = "l", xlab="interval", ylab="steps" )
```

```{r}
sortedmean <- intervals[rev(order(intervals$mean)),]
maxsteps <- sortedmean[1,2]
```
On average across all days, the `r maxsteps` interval had the maximum number of steps.  

## Imputing missing values
```{r}
missing <- sum(!complete.cases(act))
```
The dataset is missing `r missing` rows of data.  

Fill in missing values with the mean steps for that five-minute interval. 
```{r}
actcomplete <- act 
  for (i in 1:17568) {
		if (is.na(actcomplete[i,1]) == TRUE) {
			actcomplete[i,1] <- intervals$mean[intervals$int==act$interval[i]]
		}
	}
```

Below is a histogram of total steps per day for the new dataset with filled missing values. 
```{r}
daysteps <- tapply(actcomplete$steps, actcomplete$date, sum)
hist(daysteps, breaks = 50)
```

```{r}
meanday <- mean(daysteps)
medianday <- median(daysteps)
```
The mean steps per day for the new dataset is `r as.integer(meanday)`.
The median steps per day for the new data set is `r as.integer(medianday)`. 

When the missing values are filled in, the mean and median steps per day both increase.  The histograms show an overall increase in steps per day (and far fewer days with zero steps), so this is expected. 

## Are there differences in activity patterns between weekdays and weekends?

```{r}  
##create factor variable to label days as weekend/weekday
actcomplete$date <- as.POSIXct(actcomplete$date, format = "%Y-%m-%d")
actcomplete$days <- weekdays(actcomplete$date)
actcomplete$wkday <- ifelse(actcomplete$days %in% c("Saturday", "Sunday"), "weekend", "weekday")
actcomplete$wkday <- as.factor(actcomplete$wkday)
```

The plot below shows average steps taken per interval, grouped by weekday and weekend days. While weekday intervals show one large peak and several smaller peaks in average steps taken during those times, weekend days show a series of peaks that are all similar in size. 

```{r}
library(lattice)
agg2 <- aggregate(actcomplete$steps, list(day = actcomplete$wkday, int = actcomplete$interva), mean)
xyplot(x ~ int | day, agg2, type = "l", layout = c(1,2), xlab = "Interval", ylab = "Steps")
```
